Estimating contemporary migration numbers of adults for iteroparous species using kinship pairs found between populations

### 2022-02-14 latex/ms.tex
done Discussion

\begin{align}
\widehat{N_2} &= \frac{ 2 n_{\rm O, 2} n_{\rm P, 2} + 1 }{\widetilde{H}_{\rm PO, 2} + 1} ,
\label{AP-3}
\end{align}
where $\widetilde{H}_{\rm PO, 2}$ is the observed number of PO pairs found in offspring samples of size $n_{\rm O, 2}$ and parent samples of size $n_{\rm P, 2}$. The term ``$+1$'' reduces the bias especially when $\widetilde{H}_{\rm PO, 2}$ is small \cite[e.g.,][]{ecolevol2021p}, where a similar derivation of this bias correction is provided in \cite{Akita_2019}. By replacing $N_2$ by $\widehat{N_2}$ in Eqs.~\ref{AP-1} and \ref{AP-2}, we can obtain the following estimators:

\begin{align}
\widehat{M_4} &=  \frac{ N_{1} \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right) \left(2 n_{\rm O, 2} n_{\rm P, 2} + 1\right)} {2n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right) \left(\widetilde{H}_{\rm PO, 2} + 1\right)},
\label{AP-4}
\end{align}
and
\begin{align}
\widehat{m_4} &=  \frac{ \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right) \left(2 n_{\rm O, 2} n_{\rm P, 2} + 1\right)} {2n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right) \left(\widetilde{H}_{\rm PO, 2} + 1\right)}.
\label{AP-5}
\end{align}
While estimator of the migrate number (Equation~\ref{AP-4}) requires the (unknown) parent number in the population 1 ($N_{1}$), estimator of the migration rate (Equation~\ref{AP-5}) can be obtained only the observed kinship pairs.

Similar to estimating $N_2$, when PO pairs found in offspring samples and parent samples both from the population 1 in the first year are also available, we can obtain the estimator of $N_1$, written by
\begin{align}
\widehat{N_1} &= \frac{ 2 n_{\rm O, 1} n_{\rm P, 1} + 1 }{\widetilde{H}_{\rm PO, 1} + 1},
\label{AP-6}
\end{align}
where $\widetilde{H}_{\rm PO, 1}$ is the observed number of PO pairs found in offspring samples of size $n_{\rm O, 1}$ and parent samples of size $n_{\rm P, 1}$. By replacing $N_1$ by $\widehat{N_1}$ in Eq.~\ref{AP-4}, we can obtain the estimator of migration number that can be calculated only from genetic data, given by
\begin{align}
\widehat{M_5} &=  \frac{ \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right) \left(2 n_{\rm O, 1} n_{\rm P, 1} + 1\right) \left(2 n_{\rm O, 2} n_{\rm P, 2} + 1\right)} {2n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right) \left(\widetilde{H}_{\rm PO, 1} + 1\right) \left(\widetilde{H}_{\rm PO, 2} + 1\right)}.
\label{AP-7}
\end{align}

@@

To estimate the contemporary migration number or migration rate, our simulation-based results provide guidance for a sample size to ensure the required accuracy and precision, especially if the order of the number of migrant parents and parental sizes are approximately known ({\bf Table S1} in {\bf Supporting Information}). For example, when $m=0.1$ and $N_1=N_2=10^3$, sampling 10\% and 20\% parents and an equal number of offspring in two populations leads 70\% and 39\% CV of ${\hat M_5}$, respectively (case of invasive sampling). Even if there is no information about these numbers, $1/\sqrt{H_{\rm HS, bet}}$ (or $1/\sqrt{H_{\rm PO, bet}}$) provides an approximate lower bound on the CV, which can be used as an indicator of the precision of ${\hat M}$ or ${\hat m}$. Furthermore, the condition that $n>\sqrt{N_2}$ is also used as rule of thumb especially when planning a research project ($n = n_{\rm O, 1}=n_{\rm O, 2}=n_{\rm P, 2}$, see also Equations~\ref{RS-1} and \ref{RS-2}).

@@

Finally, we note some caveats in applying proposed method. Our theory for developing estimators assumes that kinships are detected without any error. There are many algorithms to detect kinship pairs from single nucleotide polymorphisms (SNPs) or short tandem repeats (STRs) \cite[e.g., ][]{Wang_2009Genetics, Huisman_2017}, although a HS pair requires many more DNA markers than a PO pair. In addition, iteroprous species potentially has several kinship types such as halfsib-uncle/nephew or halfsib-cousin, which are expected to appear frequently and should be correctly distinguished from HS pairs. It is desirable to estimate in advance how many markers are needed for kinship detections in target populations associated with simulation of a pedigree reconstruction including distant kinships (e.g., \cite{Anderson_2020}). Our proposed estimators are limited to detect parental movements on a period between breeding seasons in given populations. Thus, the estimation of fine-scale spatial-temporal movements, which is available to integrate multiple type of data \cite[]{Thorson_2021}, goes beyond the scope of this paper, although population dynamics models with a coarser spatial resolution than the spatial scale for environmental layers are often used in the assessment models, which is the target of application of the proposed method. In addition, the current theory behind the estimators does not assume desynchronized reproduction within a population (e.g., skip spawning, see also \cite{https://doi.org/10.1111/faf.12615}) and correlation between mobility and fertility. Our proposed estimators require the pre-specification of a population structure. Kinship relationships with information of a sample location potentially provide the chance to explore a plausible population structure via estimating migration number or rate between hypothetical populations, which contributes the determination of a management unit. This perspective of the study will be conducted in the future.

%  (e.g., \cite{Hastings_1993} numerically demonstrates that two populations become demographically independent when migration rate $< 0.1$)
% withinのデータはどう使う？？withinとbetweenを比較して、betweenが優位に少ないかどうかなどが構造発見につながるかもconn et al 2020 EcolEvolもチェック
%% 今回提案した方法は、ゴーストポップがあっても大丈夫

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{ACKNOWLEDGMENTS}
The author thanks Y. Tsukahara and N. Suzuki for fruitful discussions. This work was supported by JSPS KAKENHI Grant Number 19K06862 and 20H03012.

@@

\renewcommand{\arraystretch}{0.6}
\begin{table}[tb]
\begin{center}
   \caption[]{Table 2: Summary of proposed estimators for required parameters and kinship types}
    \textbf {}\\[-0mm]
    \begin{tabular}{ccccccccc} \hline
       	\\
	Estimator			& $N_1$ 		& $N_2$ 		& $n_{\rm O, 1}$	& $n_{\rm O, 2}$	& $n_{\rm P, 1}$	& $n_{\rm P, 2}$	& Required kinship type\\
	\\
	\hline
	\\
	$\widehat{M_1}$	& given 		& given 		& \checkmark		& \checkmark 		&				&				& HSP\\ 
		                						& \\
	$\widehat{M_2}$	& given		& given		& \checkmark		& 				&				& \checkmark 		& POP\\
		                						& \\
	$\widehat{M_3}$	& given		& given		& \checkmark		&  \checkmark		&				& \checkmark 		& HSP \& POP\\
		                						& \\
	$\widehat{M_4}$	& given		& estimated	& \checkmark		&  \checkmark		&				& \checkmark 		& HSP \& POP\\
		                						& \\
	$\widehat{M_5}$	& estimated	& estimated	& \checkmark		&  \checkmark		& \checkmark		& \checkmark 		& HSP \& POP\\
									& \\
	$\widehat{m_1}$	& ---			& given		& \checkmark		& \checkmark 		&				&				& HSP\\ 
									& \\
	$\widehat{m_2}$	& ---			& given		& \checkmark		& 				&				& \checkmark 		& POP\\					                			& \\
	$\widehat{m_3}$	& ---			& given 		& \checkmark		&  \checkmark		&				& \checkmark 		& HSP \& POP\\
		                						& \\
	$\widehat{m_4}$	& ---			& estimated	& \checkmark		&  \checkmark		& 				& \checkmark 		& HSP \& POP\\
		                						& \\
	\hline
    \end{tabular}
    \label{parameter} 
\end{center} 
\end{table}
\renewcommand{\arraystretch}{1}

>>>>

\begin{align}
\widehat{N_2} &= \frac{ 2 n_{\rm O, 2} n_{\rm P, 2}  }{\widetilde{H}_{\rm PO, 2} + 1} ,
\label{AP-3}
\end{align}
where $\widetilde{H}_{\rm PO, 2}$ is the observed number of PO pairs found in offspring samples of size $n_{\rm O, 2}$ and parent samples of size $n_{\rm P, 2}$. The term ``$+1$'' reduces the bias especially when $\widetilde{H}_{\rm PO, 2}$ is small \cite[e.g.,][]{ecolevol2021p}, where a similar derivation of this bias correction is provided in \cite{Akita_2019}. By replacing $N_2$ by $\widehat{N_2}$ in Eqs.~\ref{AP-1} and \ref{AP-2}, we can obtain the following estimators:

\begin{align}
\widehat{M_4} &=  \frac{ n_{\rm O, 2} n_{\rm P, 2} N_{1} \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right) } {n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right) \left(\widetilde{H}_{\rm PO, 2} + 1\right)},
\label{AP-4}
\end{align}
and
\begin{align}
\widehat{m_4} &=  \frac{ n_{\rm O, 2} n_{\rm P, 2}\left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right)} {n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right) \left(\widetilde{H}_{\rm PO, 2} + 1\right)}.
\label{AP-5}
\end{align}
While estimator of the migrate number (Equation~\ref{AP-4}) requires the (unknown) parent number in the population 1 ($N_{1}$), estimator of the migration rate (Equation~\ref{AP-5}) can be obtained only the observed kinship pairs.

Similar to estimating $N_2$, when PO pairs found in offspring samples and parent samples both from the population 1 in the first year are also available, we can obtain the estimator of $N_1$, written by
\begin{align}
\widehat{N_1} &= \frac{ 2 n_{\rm O, 1} n_{\rm P, 1} }{\widetilde{H}_{\rm PO, 1} +1},
\label{AP-6}
\end{align}
where $\widetilde{H}_{\rm PO, 1}$ is the observed number of PO pairs found in offspring samples of size $n_{\rm O, 1}$ and parent samples of size $n_{\rm P, 1}$. By replacing $N_1$ by $\widehat{N_1}$ in Eq.~\ref{AP-4}, we can obtain the estimator of migration number that can be calculated only from genetic data, given by
\begin{align}
\widehat{M_5} &=  \frac{ 2 n_{\rm O, 1} n_{\rm P, 1} n_{\rm O, 2} n_{\rm P, 2} \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right)} {n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right) \left(\widetilde{H}_{\rm PO, 1} + 1\right) \left(\widetilde{H}_{\rm PO, 2} + 1\right)}.
\label{AP-7}
\end{align}

@@

To estimate the contemporary migration number or migration rate, our simulation-based results provide guidance for a sample size to ensure the required accuracy and precision, especially if the order of the number of migrant parents and parental sizes are approximately known ({\bf Table S1} in {\bf Supporting Information}). For example, when $m=0.1$ and $N_1=N_2=10^3$, sampling 10\% and 20\% parents and an equal number of offspring in two populations leads 70\% and 39\% CV of ${\hat M_5}$, respectively (case of invasive sampling). Even if there is no information about these numbers, $1/\sqrt{H_{\rm HS, bet}}$ (or $1/\sqrt{H_{\rm PO, bet}}$) provides an approximate lower bound on the CV, which can be used as an indicator of the precision of ${\hat M}$ or ${\hat m}$. Furthermore, the condition that $n>\sqrt{N_2}$ is also used as rule of thumb especially when planning a research project ($n = n_{\rm O, 1}=n_{\rm O, 2}=n_{\rm P, 2}$, see also Equations~\ref{RS-1} and \ref{RS-2}). It should be noted that the guidance is to estimate the information about one-way migration between two populations; thus, estimating two-way migrations among two and three populations requires twice and three times the number of samples, respectively, compared to our guidance for a sample size.

@@

Finally, we note some caveats in applying proposed method. Our theory for developing estimators assumes that kinships are detected without any error. There are many algorithms to detect kinship pairs from single nucleotide polymorphisms (SNPs) or short tandem repeats (STRs) \cite[e.g., ][]{Wang_2009Genetics, Huisman_2017}, although a HS pair requires many more DNA markers than a PO pair. In addition, iteroprous species potentially has several kinship types such as halfsib-uncle/nephew or halfsib-cousin, which are expected to appear frequently and should be correctly distinguished from HS pairs. It is desirable to estimate in advance how many markers are needed for kinship detections in target populations associated with simulation of a pedigree reconstruction including distant kinships (e.g., \cite{Anderson_2020}). Our proposed estimators are limited to detect parental movements on a period between breeding seasons in given populations. Thus, the estimation of fine-scale spatial-temporal movements, which is available to integrate multiple type of data \cite[]{Thorson_2021}, goes beyond the scope of this paper, although population dynamics models with a coarser spatial resolution than the spatial scale for environmental layers are often used in the assessment models, which is the target of application of the proposed method. Our proposed estimators require the pre-specification of a population structure. Kinship relationships with information of a sample location potentially provide the chance to explore a plausible population structure via estimating migration number or rate between hypothetical populations, which contributes the determination of a management unit. In addition, the current theory behind the estimators does not assume desynchronized reproduction within a population (e.g., skip spawning, see also \cite{https://doi.org/10.1111/faf.12615}), correlation between individual mobility and fertility and non-random sampling of parents (e.g., mothers with a large number of offspring are likely to be sampled). These perspectives of the study will be conducted in the future.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{ACKNOWLEDGMENTS}
The author thanks N. Shinohara, N. Suzuki and Y. Tsukahara for fruitful discussions. This work was supported by JSPS KAKENHI Grant Number 19K06862 and 20H03012.

@@

\renewcommand{\arraystretch}{0.6}
\begin{table}[tb]
\begin{flushleft}
   \caption[]{Table 2: Summary of proposed estimators for required parameters and kinship types}
    \textbf {}\\[-0mm]
    \begin{tabular}{ccccccccc} \hline
       	\\
	Estimator			& $N_1$ 		& $N_2$ 		& $n_{\rm O, 1}$	& $n_{\rm O, 2}$	& $n_{\rm P, 1}$	& $n_{\rm P, 2}$	& Required kinship type\\
	\\
	\hline
	\\
	$\widehat{M_1}$	& given 		& given 		& \checkmark		& \checkmark 		&				&				& HSP\\ 
		                						& \\
	$\widehat{M_2}$	& given		& given		& \checkmark		& 				&				& \checkmark 		& POP\\
		                						& \\
	$\widehat{M_3}$	& given		& given		& \checkmark		&  \checkmark		&				& \checkmark 		& HSP \& POP\\
		                						& \\
	$\widehat{M_4}$	& given		& estimated	& \checkmark		&  \checkmark		&				& \checkmark 		& HSP \& POP\\
		                						& \\
	$\widehat{M_5}$	& estimated	& estimated	& \checkmark		&  \checkmark		& \checkmark		& \checkmark 		& HSP \& POP\\
									& \\
	$\widehat{m_1}$	& ---			& given		& \checkmark		& \checkmark 		&				&				& HSP\\ 
									& \\
	$\widehat{m_2}$	& ---			& given		& \checkmark		& 				&				& \checkmark 		& POP\\					                			& \\
	$\widehat{m_3}$	& ---			& given 		& \checkmark		&  \checkmark		&				& \checkmark 		& HSP \& POP\\
		                						& \\
	$\widehat{m_4}$	& ---			& estimated	& \checkmark		&  \checkmark		& 				& \checkmark 		& HSP \& POP\\
		                						& \\
	\hline
    \end{tabular}
    \\HSP and POP indicates half-sibling pair and parent-offspring pair, respectively. 
    \label{parameter} 
\end{flushleft} 
\end{table}
\renewcommand{\arraystretch}{1}
